---
title: "Reproducible Research: Peer Assessment 1"
author: Kun Xie
output: 
html_document:
keep_md: true
---


## Loading and preprocessing the data

```{r cache=TRUE}
URL="https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(URL,destfile = "rawdata.zip")
unzip("rawdata.zip")
activityData<-read.csv("activity.csv")
```

## What is mean total number of steps taken per day?

```{r}
activityDataDay<-with(activityData,aggregate(steps,by=list(date),sum,na.rm=TRUE))
hist(activityDataDay$x
     ,main="Number of total step taken in a day"
     ,xlab="Steps"
     ,ylab="Days")
meanstep<-format(mean(activityDataDay$x),digits = 1)
medianstep<-format(median(activityDataDay$x),digits = 1)
```

The mean and median number of steps taken per day, assuming NA indicate zero steps, is `r meanstep` and `r medianstep` respectively


## What is the average daily activity pattern?

```{r}
activityDataInt<-with(activityData,aggregate(steps,list(interval),FUN=mean,na.rm=TRUE))
with(activityDataInt,plot(Group.1,x,type='l',xlab="Internvals",ylab="Average step Count"))
maxStep<-max(activityDataInt$x)
maxInterval<-activityDataInt[activityDataInt$x==maxStep,]$Group.1
```

The inverval with max amount of average step taken is at `r maxInterval` minute with a average steps taken of `r maxStep`


## Imputing missing values

```{r}
NACount<-sum(is.na(activityData$steps))
#replace the NA value with mean value of that time interval
activityDataImp<-merge(activityData,activityDataInt,by.x="interval",by.y="Group.1")
activityDataImp[is.na(activityDataImp$steps),]$steps<-activityDataImp[is.na(activityDataImp$steps),]$x
head(activityDataImp)
#perform the same analysis as before on imputed dataframe
activityDataDay<-with(activityDataImp,aggregate(steps,by=list(date),sum,na.rm=TRUE))
hist(activityDataDay$x
     ,main="Number of total step taken in a day (Imputed)"
     ,xlab="Steps"
     ,ylab="Days")
meanstepImp<-format(mean(activityDataDay$x),digits = 1)
medianstepImp<-format(median(activityDataDay$x),digits = 1)
```

1. There are `r NACount` NA in original dataset.
2. We replace the N/A in original dataset with mean of the specific interval across all days before imputing
3. The new mean is `r meanstepImp` and median is `r medianstepImp`
4. Compare to orignal data, both the mean and median is larger after imputing


## Are there differences in activity patterns between weekdays and weekends?

```{r}
#subset the dataset based on weekdays
activityDataImp$weekdays<-weekdays(as.Date(activityDataImp$date))
activityDataImp[activityDataImp$weekdays!=c("Saturday","Sunday"),]$weekdays<-"Weekday"
activityDataImp[activityDataImp$weekdays==c("Saturday","Sunday"),]$weekdays<-"Weekend"
activityDataInt<-with(activityDataImp,aggregate(steps,by=list(interval,weekdays),FUN=mean,na.rm=TRUE))

#load lattice for panel plot, and plot the dataset
library(lattice)
xyplot(x~Group.1|Group.2,data=activityDataInt,type="l",layout=c(1,2),
       xlab="Interval",
       ylab="Average Number of Steps")

```

Based on the above pattern, it would appear that during day time, people are generally more active at weekend than weekdays.






